Combinatorial Identities from Lagrange's Interpolation Polynomial

### 2020-11-06 spectral radius of matrix new method/thesis/our_result.tex
1106 update with success labels but unknown packages

\usepackage{fontspec}   %加這個就可以設定字體
\usepackage{xeCJK}       %讓中英文字體分開設置
%\usepackage{times}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{tabularx,array}
\usepackage{pgf,tikz}
\usepackage{blkarray} %line 80-92 Q formula matrix index
%\usepackage{colortbl}
\usepackage{enumitem}% enumerate labels   roman

@@

\chapter{Our result}
We will prove the main Theorems introduced in the introduction. First, we need some natations.
 Throughout fix $k \in [n-1] := \{1,2,...,n-1\}$. Let $E_{kn}$ denote the $n\times n$ binary matrix with an unique $1$ appearing in the  position $(k,n)$. We will apply the Theorem 1.7 with $P=I$ and

\begin{equation}
  \label{Q_1}
Q=I+E_{kn}=\begin{blockarray}{cccccc}
%a & b & c & d & e \\
 &  &  & &  \\
\begin{block}{(ccccc)c}

  1 &   &   &   &   &  \\
    & 1 &   &   &   &  \\
    &   & \ddots &   & 1 & k \\
    &   &   & 1 &   &  \\
  0 &   &   &   & 1 &  \\
\end{block}
\end{blockarray}
\end{equation}

@@

\begin{lem}\label{lem:rt_vec}
If $u=(u_1, u_2, \ldots, u_n)^T$ and $v'=(v'_1, v'_2, \ldots, v'_n):=Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$, then
\begin{enumerate}[label=(\Roman*)]
\item \label{lem:rt_vec:en1}$v'$ is $k$-rooted  if and only if  $u$ is nonnegative;
\item $u_k>0$ if and only if $v'_k>v'_n$.
\end{enumerate}
\qed
\end{lem}

\begin{thm} \label{thm_main}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}[label=(\Roman*)]
\item \label{thm_main:condition_i} $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
\item \label{thm_main:condition_ii}there exists a $k$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
\item there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
\item $v^Tv'>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
Moreover, $\lambda=\lambda'$
if and only if
\begin{enumerate}[label=(\alph*)]
\item  \label{thm_main:equ_cond_a} $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i\not=0$ and $v'_n\not=0;$
\item \label{thm_main:equ_cond_b} $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1, j \neq k $with $v_i\ne 0 $;
\item \label{thm_main:equ_cond_c} $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ and $ v'_{k}>v'_n$ 
\end{enumerate} \qed
\end{thm}

@@

\begin{defn}\label{m_rooted}
A  matrix $C'=(c'_{ij})$ is called $k-rooted$  if its  columns and its $(k, n)-sum$ vector are all $k$-rooted except the last column of $C'$.
\end{defn}

The following theorem shows that a $k$-rooted matrix has a $k$-rooted eigenvector.

\begin{lem} \label{lma_m_rooted}
Let $C'=(c'_{ij})$ be an $n\times n$ nonnegative matrix. Then the following (i)-(iii) hold.
    \begin{enumerate}[label=(\Roman*)]
        \item $C'$ is a $k$-rooted matrix, if and only if, $Q^{-1}C'Q$ is nonnegative.
        \item Assume that $C'$ is $k$-rooted and let $u$ be a nonnegative eigenvector of $Q^{-1}C'Q$
         for $\rho(C')$. Then  $C'$ has a $k$-rooted eigenvector $v'=Qu$ for $\rho(C')$. 
        \item $\rho(C')$ = $\rho(Q^{-1}C'Q)$
    \end{enumerate}
\end{lem}

>>>>

\usepackage{fontspec}   %加這個就可以設定字體
\usepackage{xeCJK}       %讓中英文字體分開設置
%\usepackage{times}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{ulem}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{tabularx,array}
\usepackage{pgf,tikz}
\usepackage{blkarray} %begin{equation}   Q=I+E_{kn} : Q formula matrix index
%\usepackage{colortbl}
\usepackage{enumitem}% enumerate labels   roman

@@

\chapter{Our result}

\begin{rem}[theorem in previous slide]\label{pre_thm}
  Let $C=(c_{ij})$, $C'=(c'_{ij})$, $P$ and $Q$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}[label=(\roman*)]
  \item \label{pre_thm_em1}  $PCQ\leq PC'Q$;   %
  \item \label{pre_thm_em2}  there exist a nonnegative column vector $u=(u_1, u_2, \ldots, u_n)^T$  and a
  scalar $\lambda'\in \mathbb{R}$ such that $\lambda'$ is an eigenvalue of $C'$ with
  associated eigenvector $Qu$;
  \item \label{pre_thm_em3}  there exist a nonnegative row vector $v^T=(v_1, v_2, \ldots, v_n)$  and a scalar
  $\lambda\in \mathbb{R}$such that $\lambda$ is an eigenvalue of $C$ with associated  left
  eigenvector $v^TP$; and
  \item \label{pre_thm_em4} $v^TPQu>0.$
\end{enumerate}
  Then $\lambda\leq \lambda'$. Moreover, $\lambda=\lambda'$if and only if
  \begin{equation}\label{pre0}
      (PC'Q)_{ij}=(PCQ)_{ij}\qquad \hbox{for~}1\leq i, j\leq n \hbox{~with~} v_i\ne 0 \hbox{~and~} u_j\ne 0.
  \end{equation}
\end{rem}

\begin{pof}
  Multiplying the nonnegative vector $u$ in theorem~\ref{pre_thm} assumption
  ~\ref{pre_thm_em1}, where $Qu$ is eigenvector of $C'$,  to the right of both terms of
  ~\ref{pre_thm_em1},
  \begin{equation}\label{pre1}
      PCQu\leq PC'Qu=\lambda'PQu.
  \end{equation}
  Multiplying the nonnegative left eigenvector $v^T$ of $C$ for $\lambda$ in assumption
  ~\ref{pre_thm_em3} to the left of all terms  in (\ref{pre1}), where $v^TP$ is
  left eigenvector of $C$ for $\lambda$, thus we have
  \begin{equation}\label{pre2}
      \lambda v^TPQu=v^TPCQu\leq v^TPC'Qu=\lambda' v^TPQu.
  \end{equation}
      Now delete the positive term $v^TPQu$ by assumption \ref{pre_thm_em4} to obtain
       $\lambda\leq \lambda'$ and finish the proof of the first part.
  Assume that $\lambda=\lambda'$, so the inequality in (\ref{pre2}) is an equality.
  Especially $(PCQu)_i=(PC'Qu)_i$ for any $i$ with $v_i\not=0.$ Hence,
  $(PCQ)_{ij}=(PC'Q)_{ij}$ for any $i$ with $v_i\not=0$ and any $j$ with
  $u_j\not=0.$ Conversely, (\ref{pre0}) implies $$v^TPCQu=\sum_{i,j} v_i(PCQ)_{ij}u_j=
  \sum_{i,j} v_i(PC'Q)_{ij}u_j=v^TPC'Qu,$$ so $\lambda=\lambda'$ by (\ref{pre2}).

We will prove the main Theorems introduced in the introduction. First, we need some natations.
 Throughout fix $k \in [n-1] := \{1,2,...,n-1\}$. Let $E_{kn}$ denote the $n\times n$ binary matrix with an unique $1$ appearing in the  position $(k,n)$. We will apply the Theorem 1.7 with $P=I$ and

\begin{equation} 
Q=I+E_{kn}=\begin{blockarray}{cccccc} 
%a & b & c & d & e \\
 &  &  & & & \\  
\begin{block}{(ccccc)c}
	% \label{Q_1}
  1 &   &   &   &   &  \\
    & 1 &   &   &   &  \\
    &   & \ddots &   & 1 & k \\
    &   &   & 1 &   &  \\
  0 &   &   &   & 1 &  \\
\end{block}
\end{blockarray}
\end{equation}  \label{Q_1}

@@

\begin{lem}\label{lem:rt_vec}
If $u=(u_1, u_2, \ldots, u_n)^T$ and $v'=(v'_1, v'_2, \ldots, v'_n):=Qu=(u_1,\ldots, u_{k-1},u_k+u_n, u_{k+1}, \ldots,  u_n)^T$, then
\begin{enumerate}[label=(\roman*)]
\item \label{lem:rt_vec:en1}$v'$ is $k$-rooted  if and only if  $u$ is nonnegative;
\item $u_k>0$ if and only if $v'_k>v'_n$.
\end{enumerate}
\qed
\end{lem}

\begin{thm} \label{thm_main}
    Let $C=(c_{ij})$, $C'=(c'_{ij})$ be  $n\times n$ matrices.
Assume that
\begin{enumerate}[label=(\roman*)]
\item \label{thm_main:condition_i} $C[[n]|[n-1]]\leq C'[[n]|[n-1]]$ and $c_{ik}+c_{in}\leq c'_{ik}+c'_{in}$ for all $1\leq i\leq n$;
\item \label{thm_main:condition_ii}there exists a $k$-rooted vector $v'=(v'_1, v'_2, \ldots, v'_n)^T$ and a scalar $\lambda'\in \mathbb{R}$
such that $\lambda'$ is an eigenvalue of $C'$ with associated eigenvector $v'$;
\item \label{thm_main:condition_iii} there exists a nonnegative vector $v^T=(v_1, v_2, \ldots, v_n)$ and a scalar $\lambda\in \mathbb{R}$ such that $\lambda$ is an eigenvalue of $C$ with associated left eigenvector $v^T$;
\item \label{thm_main:condition_iv} $v^Tv'>0.$
\end{enumerate}
 Then $\lambda\leq \lambda'$.
Moreover, $\lambda=\lambda'$
if and only if
\begin{enumerate}[label=(\alph*)]
\item  \label{thm_main:equ_cond_a} $c_{ik}+c_{in}=c'_{ik}+c'_{in} \qquad$  for $1\leq i\leq n$ with $v_i\not=0$ and $v'_n\not=0;$
\item \label{thm_main:equ_cond_b} $c'_{ij}=c_{ij}\qquad $for $1\leq i\leq n,~1\leq j\leq n-1, j \neq k $with $v_i\ne 0 $;
\item \label{thm_main:equ_cond_c} $c'_{ik}=c_{ik} \qquad $  for $1\leq i \leq n$ and $ v'_{k}>v'_n$ 
\end{enumerate} \qed
\end{thm}

@@

\begin{defn}\label{m_rooted}
A  matrix $C'=(c'_{ij})$ is called $k-rooted$  if its  columns and its $(k, n)-sum$ vector are all $k-rooted$ except the last column of $C'$.
\end{defn}

The following theorem shows that a $k$-rooted matrix has a $k$-rooted eigenvector. 
\begin{lem} \label{lma_m_rooted}
Let $C'=(c'_{ij})$ be an $n\times n$ nonnegative matrix. Then the following (i)-(iii) hold.
    \begin{enumerate}[label=(\roman*)]
        \item $C'$ is a $k$-rooted matrix, if and only if, $Q^{-1}C'Q$ is nonnegative.
        \item Assume that $C'$ is $k$-rooted and let $u$ be a nonnegative eigenvector of $Q^{-1}C'Q$
         for $\rho(C')$. Then  $C'$ has a $k$-rooted eigenvector $v'=Qu$ for $\rho(C')$. 
        \item $\rho(C')$ = $\rho(Q^{-1}C'Q)$
    \end{enumerate}
\end{lem}
